---
title: "introduction-to-text-analysis-in-r-chapter-3"
author: "Xiaotian Sun"
date: "09/02/2020"
output: html_document
---

## 3. Sentiment Analysis

##### 3.1 Sentiment dictionaries (video)

- The tidytext pacakage includes four distinct sentiment dictionaries that can be accessed with the get_sentiments() function. The only argument is the name of the dictionary.

- dictionary: `bing`, `afinn`, `loughran`, `nrc`.

##### 3.2 Counting the NRC sentiments 

The fourth dictionary included with the tidytext package is the `nrc` dictionary. Let's start our exploration with sentiment counts.

***Instruction :***

- I usually do this for you, but start with loading the `tidyverse` and `tidytext` packages.

- Count the number of words associated with each sentiment in `nrc`.

- Arrange the counts in descending order.

```{r}
# Load the tidyverse and tidytext packages
library(tidyverse)
library(tidytext)

# Count the number of words associated with each sentiment in nrc
get_sentiments("nrc") %>% 
  count(sentiment) %>% 
  # Arrange the counts in descending order
  arrange(desc(n))
```

##### 3.3 Visualizing the NRC sentiments 

We've seen how visualizations can give us a better idea of patterns in data than counts alone. Let's visualize the sentiments from the `nrc` dictionary. I've loaded the `tidyverse` and `tidytext` packages for you already.

***Instruction :***

- Extract the `nrc` dictionary, count the sentiments and reorder them by count to create a new factor column, `sentiment2`.

- Visualize `sentiment_counts` using the new sentiment factor column.

- Change the title to "Sentiment Counts in NRC", x-axis to "Sentiment", and y-axis to "Counts".

```{r}
# Pull in the nrc dictionary, count the sentiments and reorder them by count
sentiment_counts <- get_sentiments("nrc") %>% 
  count(sentiment) %>% 
  mutate(sentiment2 = fct_reorder(sentiment, n))

# Visualize sentiment_counts using the new sentiment factor column
ggplot(sentiment_counts, aes(x = sentiment2, y = n)) +
  geom_col() +
  coord_flip() +
  # Change the title to "Sentiment Counts in NRC", x-axis to "Sentiment", and y-axis to "Counts"
  labs(
    title = "Sentiment Counts in NRC",
    x = "Sentiment",
    y = "Counts"
  )
```

##### 3.4 Appending dictionaries (video)

- We need to first append the sentiment dictionary to our tokenized and cleaned text data.

- inner_join()

- We use `scales = "free"` to set both axes free with "free".

##### 3.5 Counting sentiment 

The `tidy_twitter` dataset has been loaded for you. Let's see what sort of sentiments are most prevalent in our Twitter data.

```{r echo=FALSE}
twitter_data <- readRDS("ch_1_twitter_data.rds")

tidy_twitter <- twitter_data %>% 
  # Tokenize the twitter data
  unnest_tokens(word, tweet_text) %>% 
  # Remove stop words
  anti_join(stop_words2)
```


***Instruction :***

- Join `tidy_twitter` and the NRC sentiment dictionary.

- Count the sentiments in `sentiment_twitter`.

- Arrange the sentiment counts in descending order.

```{r}
# Join tidy_twitter and the NRC sentiment dictionary
sentiment_twitter <- tidy_twitter %>% 
  inner_join(get_sentiments("nrc"))

# Count the sentiments in sentiment_twitter
sentiment_twitter %>% 
  count(sentiment) %>% 
  # Arrange the sentiment counts in descending order
  arrange(desc(n))
```

##### 3.6 Visualizing sentiment

Let's explore which words are associated with each sentiment in our Twitter data.

***Instruction 1:***

- Inner join `tidy_twitter` to the NRC dictionary and filter for positive, fear, and trust.

- Count by word and sentiment and keep only the top 10 of each sentiment.

- Create a factor called `word2` that has each word ordered by the count.

```{r}
word_counts <- tidy_twitter %>% 
  # Append the NRC dictionary and filter for positive, fear, and trust
  inner_join(get_sentiments("nrc")) %>% 
  filter(sentiment %in% c("positive", "fear", "trust")) %>%
  # Count by word and sentiment and keep the top 10 of each
  count(word, sentiment) %>% 
  group_by(sentiment) %>% 
  top_n(10, n) %>% 
  ungroup() %>% 
  # Create a factor called word2 that has each word ordered by the count
  mutate(word2 = fct_reorder(word, n))
```

***Instruction 2:***

- Create a bar plot of the word counts colored by sentiment.

- Create a separate facet for each sentiment with free axes.

- Title the plot "Sentiment Word Counts" with "Words" for the x-axis.

```{r}
# Create a bar plot out of the word counts colored by sentiment
ggplot(word_counts, aes(x = word2, y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  # Create a separate facet for each sentiment with free axes
  facet_wrap(~ sentiment, scales ="free") +
  coord_flip() +
  # Title the plot "Sentiment Word Counts" with "Words" for the x-axis
  labs(
    title = "Sentiment Word Counts",
    x = "Words"
  )
```

##### 3.7 Improving sentiment analysis (video)

- spread() 

##### 3.8 Practicing reshaping data 

The `spread()` verb allows us to quickly reshape or stack and transpose our data, making it easier to `mutate()`.

***Instruction :***

- Append `tidy_twitter` to the NRC sentiment dictionary.

- Count by complaint label and sentiment.

- Spread the sentiment and count columns.

```{r}
tidy_twitter %>% 
  # Append the NRC sentiment dictionary
  inner_join(get_sentiments("nrc")) %>% 
  # Count by complaint label and sentiment
  count(complaint_label, sentiment) %>% 
  # Spread the sentiment and count columns
  spread(sentiment, n)
```

##### 3.9 Practicing with grouped summaries 

We can use `spread()` in association with the output of grouped summaries as well.

***Instruction :***

- Append `tidy_twitter` to the afinn sentiment dictionary.

- Group by both complaint label and whether or not the user is verified.

- Summarize the data to create a new column, `aggregate_value`, which contains the sum of `value`.

- Spread the `complaint_label` and `aggregate_value` columns.

```{r}
tidy_twitter %>% 
  # Append the afinn sentiment dictionary
  inner_join(get_sentiments("afinn")) %>% 
  # Group by both complaint label and whether or not the user is verified
  group_by(complaint_label, usr_verified) %>% 
  # Summarize the data with an aggregate_value = sum(value)
  summarise(aggregate_value = sum(value)) %>% 
  # Spread the complaint_label and aggregate_value columns
  spread(complaint_label, aggregate_value) %>% 
  mutate(overall_sentiment = Complaint + `Non-Complaint`)
```

##### 3.10 Visualizing sentiment by complaint type

Now let's see whether or not complaints really are more negative, on average.

***Instruction 1:***

- Append `tidy_twitter` to the bing sentiment dictionary.

- Count by complaint label and sentiment.

- Spread the sentiment and count columns.

- Add a new column, `overall_sentiment`, as `positive - negative`.

```{r}
sentiment_twitter <- tidy_twitter %>% 
  # Append the bing sentiment dictionary
  inner_join(get_sentiments("bing")) %>% 
  # Count by complaint label and sentiment
  count(complaint_label, sentiment) %>% 
  # Spread the sentiment and count columns
  spread(sentiment, n) %>% 
  # Compute overall_sentiment = positive - negative
  mutate(overall_sentiment = positive - negative)
```

***Instruction 2:***

- Create a bar plot of overall sentiment by complaint label, colored by complaint label (as a factor).

- Title the plot "Overall Sentiment by Complaint Label" with the subtitle "Airline Twitter Data".

```{r}
# Create a bar plot out of overall sentiment by complaint level, colored by a complaint label factor
ggplot(
  sentiment_twitter, 
  aes(x = complaint_label, y = overall_sentiment, fill = as.factor(complaint_label))
) +
  geom_col(show.legend = FALSE) +
  coord_flip() + 
  # Title the plot "Overall Sentiment by Complaint Type," with an "Airline Twitter Data" subtitle
  labs(
    title = "Overall Sentiment by Complaint Type",
    subtitle = "Airline Twitter Data"
  )
```
