---
title: "chapter-5-case-studies"
author: "Xiaotian Sun"
date: "14/03/2020"
output: html_document
---

## 5. Case Studies

##### 5.1 A case study, reading a play (video)

##### 5.2 Getting the play into R 

We've already downloaded the play and put the text file in your workspace. Your first step is to read the play into R using `stri_read_lines()`.

You should take a look at the original text file: `importance-of-being-earnest.txt`

You'll see there is some foreword and afterword text that Project Gutenberg has added. You'll want to remove that, and then split the play into the introduction (the list of characters, scenes, etc.) and the main body.

```{r echo=FALSE}
library(tidyverse)
library(stringr)
library(stringi)
library(rebus)
earnest_file <- "importance-of-being-earnest.txt"
```

***Instruction 1:***

Use `stri_read_lines()` to read the play from `earnest_file` into `earnest`.

```{r}
# Read play in using stri_read_lines()
earnest <- stri_read_lines(earnest_file)
```

***Instruction 2:***

- Find the lines that end the foreword and start of afterword by detecting the patterns `"START OF THE PROJECT"` and `"END OF THE PROJECT"`.

- Use the `start` and `end` positions to subset the play to the lines between `(start + 1)` and `(end - 1)`.

```{r}
# Read play in using stri_read_lines()
earnest <- stri_read_lines(earnest_file)

# Detect start and end lines
start <- which(str_detect(earnest, fixed("START OF THE PROJECT")))
end <- which(str_detect(earnest, fixed("END OF THE PROJECT")))

# Get rid of gutenberg intro text
earnest_sub  <- earnest[(start + 1):(end - 1)]
```

***Instruction 3:***

- Find the lines which start the first act by looking for the pattern `"FIRST ACT"`.

- Build an index that captures the `intro_line_index` with `1:(lines_start - 1)`.

- Use `intro_line_index` to extract the intro text and `-intro_line_index` to extract the play text.

```{r}
# Read play in using stri_read_lines()
earnest <- stri_read_lines(earnest_file)

# Detect start and end lines
start <- str_which(earnest, fixed("START OF THE PROJECT"))
end <- str_which(earnest, fixed("END OF THE PROJECT"))

# Get rid of gutenberg intro text
earnest_sub  <- earnest[(start + 1):(end - 1)]

# Detect first act

lines_start <- which(str_detect(earnest_sub, fixed("FIRST ACT")))

# Set up index
intro_line_index <- 1:(lines_start - 1)

# Split play into intro and play
intro_text <- earnest_sub[intro_line_index]
play_text <- earnest_sub[-intro_line_index]
```

***Instruction 4:***

Look at the first twenty lines using `writeLines()` on a subset of `play_text`.

```{r}
# Read play in using stri_read_lines()
earnest <- stri_read_lines(earnest_file)

# Detect start and end lines
start <- str_which(earnest, fixed("START OF THE PROJECT"))
end <- str_which(earnest, fixed("END OF THE PROJECT"))

# Get rid of gutenberg intro text
earnest_sub  <- earnest[(start + 1):(end - 1)]

# Detect first act
lines_start <- str_which(earnest_sub, fixed("FIRST ACT"))

# Set up index
intro_line_index <- 1:(lines_start - 1)

# Split play into intro and play
intro_text <- earnest_sub[intro_line_index]
play_text <- earnest_sub[-intro_line_index]

# Take a look at the first 20 lines
writeLines(play_text[1:20])
```

##### 5.3 Identifying the lines, take 1 

The first thing you might notice when you look at your vector `play_text` is there are lots of empty lines. They don't really affect your task so you might want to remove them. The easiest way to find empty strings is to use the `stringi` function `stri_isempty()`, which returns a logical you can use to subset the not-empty strings:

># Get rid of empty strings
empty <- stri_isempty(play_text)
play_lines <- play_text[!empty]

So, how are you going to find the elements that indicate a character starts their line? Consider the following lines

>> play_lines[10:15]
[1] "Algernon.  I'm sorry for that, for your sake.  I don't play"             
[2] "accurately--any one can play accurately--but I play with wonderful"      
[3] "expression.  As far as the piano is concerned, sentiment is my forte.  I"
[4] "keep science for Life."                                                  
[5] "Lane.  Yes, sir."                                                        
[6] "Algernon.  And, speaking of the science of Life, have you got the"

The first line is for `Algernon`, the next three strings are continuations of that line, then line 5 is for `Lane` and line `6` for `Algernon`.

How about looking for lines that start with a word followed by a `.`?

`play_lines`, containing the lines of the play as a character vector, has been pre-defined.

***Instruction 1:***

- Build a pattern that matches the start of the line, followed by one or more word characters, then a period.

- Use your pattern with `str_view()` to see the lines that matched, and those that didn't match. Do you see any problems?

```{r}
play_lines <- play_text
```

```{r}
# Pattern for start word then .
pattern_1 <- START %R% one_or_more(WRD) %R% DOT

# Test pattern_1
str_view(play_lines, pattern = pattern_1, match = TRUE) #to see matched lines
str_view(play_lines, pattern = pattern_1, match = FALSE) 
```

***Instruction 2:***

- Try a more specific pattern: the start of the line, a capital letter, followed by one or more word characters, then a full stop.

- As before, view the matched lines, by calling `str_view()` with `match = TRUE` and `match = FALSE`.

```{r}
# Pattern for start, capital, word then .
pattern_2 <- START %R% ascii_upper() %R% one_or_more(WRD) %R% DOT

# Test pattern_2
str_view(play_lines, pattern_2, match = TRUE)
str_view(play_lines, pattern_2, match = FALSE)
```

***Instruction 3:***

- Use `str_subset()` to get just the lines that match our pattern.

- Use `str_extract()` to get the part of the string that matched our pattern in `lines`.

- Use `unique()` on `who` to see if we picked out the characters.

```{r}
# Pattern from last step
pattern_2 <- START %R% ascii_upper() %R% one_or_more(WRD) %R% DOT

# Get subset of lines that match
lines <- str_subset(play_lines, pattern = pattern_2)

# Extract match from lines
who <- str_extract(lines, pattern = pattern_2)

# Let's see what we have
unique(who)
```

##### 5.4 Identifying the lines, take 2 

The pattern "starts with a capital letter, has some other characters then a full stop" wasn't specific enough. You ended up matching lines that started with things like `University.`, `July.`, `London.`, and you missed characters like `Lady Bracknell` and `Miss Prism`.

Let's take a different approach. You know the characters names from the play introduction. So, try specifically looking for lines that start with their names. You'll find the `or1()` function from the `rebus` package helpful. It specifies alternatives but rather than each alternative being an argument like in `or()`, you can pass in a vector of alternatives.

We've created the `characters` vector for you with all the characters names.

***Instruction 1:***

- Create a pattern that matches the start of the string, followed by one of the characters names using `or1()`, followed by a `.`.

- Test your pattern.

  - View matches of `pattern_3` on the `play_lines` character vector.
  
  - View non-matches of the same.
  
  - Does it look like it worked?

```{r}
# Create vector of characters
characters <- c("Algernon", "Jack", "Lane", "Cecily", "Gwendolen", "Chasuble", 
  "Merriman", "Lady Bracknell", "Miss Prism")

# Match start, then character name, then .
pattern_3 <- START %R% or1(characters) %R% DOT

# View matches of pattern_3
str_view(play_lines, pattern = pattern_3, match = TRUE)
  
# View non-matches of pattern_3
str_view(play_lines, pattern = pattern_3, match = FALSE)
```

***Instruction 2:***

- Pull out the lines that match with `str_subset()`.

- Pull out the parts that matched with `str_extract()`.

- Take a look at the `unique()` characters you found.

- Finally, tally up how many lines each character had, but using `table()` on `who`.

```{r}
# Variables from previous step
characters <- c("Algernon", "Jack", "Lane", "Cecily", "Gwendolen", "Chasuble", 
  "Merriman", "Lady Bracknell", "Miss Prism")
pattern_3 <- START %R% or1(characters) %R% DOT

# Pull out matches
lines <- str_subset(play_lines, pattern = pattern_3)

# Extract match from lines
who <- str_extract(lines, pattern = pattern_3)

# Let's see what we have
unique(who)

# Count lines per character
table(who)
```

##### 5.5 A case study on case (video)

##### 5.6 Changing case to ease matching 

A simple solution to working with strings in mixed case, is to simply transform them into all lower or all upper case. Depending on your choice, you can then specify your pattern in the same case.

For example, while looking for `"cat"` finds no matches in the following string,

>x <- c("Cat", "CAT", "cAt") 
str_view(x, "cat")

transforming the string to lower case first ensures all variations match.

> str_view(str_to_lower(x), "cat")

See if you can find the catcidents that also involved dogs. You'll see a new `rebus` function called `whole_word()`. The argument to `whole_word()` will only match if it occurs as a word on its own, for example `whole_word("cat")` will match `cat` in `"The cat"` and `"cat."` but not in `"caterpillar"`.

A character vector of cat-related accidents has been pre-defined in your workspace as `catcidents`.

***Instruction 1:***

- Create a pattern to match `"DOG"` as a whole word using `whole_word()`, assigning the result to `whole_dog_pattern`.

- View the matches to `whole_dog_pattern` in the original mixed case `catcidents`.

```{r}
catcidents <- readRDS("catcidents.rds")
```

```{r}
# catcidents has been pre-defined
head(catcidents)

# Construct pattern of DOG in boundaries
whole_dog_pattern <- whole_word("DOG")

# View matches to word "DOG"
str_view(catcidents, pattern = whole_dog_pattern, match = TRUE)
```

***Instruction 2:***

- Create `catcidents_upper` by transforming all the strings in `catcidents` to upper case.

- Take another look at the matches to `whole_dog_pattern`, this time using `catcidents_upper` as the input.

```{r}
# From previous step
whole_dog_pattern <- whole_word("DOG")

# Transform catcidents to upper case
catcidents_upper <- str_to_upper(catcidents)

# View matches to word "DOG" again
str_view(catcidents_upper, pattern = whole_dog_pattern, match = TRUE)
```

***Instruction 3:***

If you need to retain the original mixed case strings, you can use` str_detect()` on the transformed strings to subset the original strings.

- Try it by creating `has_dog` from calling `str_detect()` on `catcidents_upper` with the upper case pattern.

- Use `has_dog` and square brackets to subset `catcidents`.

```{r}
# From previous steps
whole_dog_pattern <- whole_word("DOG")
catcidents_upper <- str_to_upper(catcidents)

# Which strings match?
has_dog <- str_detect(catcidents_upper, pattern = whole_dog_pattern)

# Pull out matching strings in original 
catcidents[has_dog]
```

##### 5.7 Ignoring case when matching 

Rather than transforming the input strings, another approach is to specify that the matching should be case insensitive. This is one of the options to the `stringr` `regex()` function.

Take our previous example,

> x <- c("Cat", "CAT", "cAt") 
str_view(x, "cat")

To match the pattern `cat` in a case insensitive way, we wrap our pattern in `regex()` and specify the argument `ignore_case = TRUE`,

> str_view(x, regex("cat", ignore_case = TRUE))

Notice that the matches retain their original case and any variant of `cat` matches.

Try it out to find the catcidents that involved tripping.

***Instruction 1:***

First view the matches of `catcidents` to the pattern `"TRIP"`. Notice how you only match those that are `TRIP` all in upper case.

Construct a case-insensitive regex to `"TRIP"` by calling `regex()` with `ignore_case = TRUE`. Assign the result to `trip_pattern`.

Repeat your viewing of catcident trips, this time using the case insensitive `trip_pattern`. You should get a few more hits.

```{r}
# View matches to "TRIP"
str_view(catcidents, pattern = "TRIP", match = TRUE)

# Construct case insensitive pattern
trip_pattern <- regex("TRIP", ignore_case = TRUE)

# View case insensitive matches to "TRIP"
str_view(catcidents, pattern = trip_pattern, match = TRUE)
```

***Instruction 2:***

- Extract only the catcidents that matched `trip_pattern` by using `str_subset()`. Assign the result to `trip`.

- Use `str_extract()` on `trip` with `trip_pattern` to pull out the matching parts.

```{r}
# From previous step
trip_pattern <- regex("TRIP", ignore_case = TRUE)

# Get subset of matches
trip <- str_subset(catcidents, pattern = trip_pattern)

# Extract matches
str_extract(trip, pattern = trip_pattern)
```

##### 5.8 Fixing case problems 

Finally, you might want to transform strings to a common case. You've seen you can use `str_to_upper()` and `str_to_lower()`, but there is also `str_to_title()` which transforms to title case, in which every word starts with a capital letter.

This is another situation where `stringi` functions offer slightly more functionality than the `stringr` functions. The `stringi` function `stri_trans_totitle()` allows a specification of the `type` which, by default, is `"word"`, resulting in title case, but can also be `"sentence"` to give sentence case: only the first word in each sentence is capitalized.

Try outputting the `catcidents` in a consistent case.

***Instruction:***

A character vector of cat-related accidents has been pre-defined in your workspace as `catcidents`.

- Store the first five elements of `catcidents` as `cat5`.

- Use `writeLines()` to display `cat5`.

- Repeat but now pass `cat5` transformed to title case with `str_to_title()`.

- Try again using `stri_trans_totitle()` instead. This should be identical to `str_to_title()`.

- Finally, display the first 5 elements of `cat5` transformed to sentence case, by passing the `type` argument to `stri_trans_totitle()`.

```{r}
library(stringi)

# Get first five catcidents
cat5 <- catcidents[1:5]

# Take a look at original
writeLines(cat5)

# Transform to title case
writeLines(str_to_title(cat5))

# Transform to title case with stringi
writeLines(stri_trans_totitle(cat5)) #same

# Transform to sentence case with stringi
writeLines(stri_trans_totitle(cat5, type = "sentence"))
```


##### 5.9 Wrapping up 

##### 5.10 An interview with Hadley Wickham