---
title: "chapter-3-pattern-matching-with-regular-expressions"
author: "Xiaotian Sun"
date: "14/03/2020"
output: html_document
---

## 3. Pattern Matching with Regular Expressions

##### 3.1 Regular expressions (video)

regular expressions as a pattern argument

##### 3.2 Matching the start or end of the string 

`rebus` provides `START` and `END` shortcuts to specify regular expressions that match the start and end of the string. These are also known as anchors. You can try it out just by typing

> START

You'll see the output `<regex>`. The `<regex>` denotes this is a special `regex` object and it has the value `^`. `^` is the character used in the regular expression language to denote the start of a string.

The special operator provided by `rebus`, `%R%` allows you to compose complicated regular expressions from simple pieces. When you are reading `rebus` code, think of `%R%` as "then". For example, you could combine `START` with `c`,

> START %R% "c"

to match the pattern "the start of string then a `c`", or in other words: strings that start with `c`. In `rebus`, if you want to match a specific character, or a specific sequence of characters, you simply specify them as a string, e.g. surround them with `"`.

***Instruction 1:***

- Which character is used to match the end of a string? Print `END` to find out.

- `str_view()` from `stringr` is really helpful for testing your patterns. Run this line of code to see the matches in `x` to the pattern `START %R% "c"`.

```{r echo=FALSE}
library(stringr)
library(rebus)
```

```{r}
# Some strings to practice with
x <- c("cat", "coat", "scotland", "tic toc")

# Print END
END

# Run me
str_view(x, pattern = START %R% "c")
```

***Instruction 2:***

Provide the pattern argument to match strings that start in `"co"`.

```{r}
# Match the strings that start with "co" 
str_view(x, pattern = START %R% "co")
```

***Instruction 3:***

Provide the pattern argument to match strings that end in `"at"`. Think about this as matching an `"at"` followed by the end of the string.

```{r}
# Match the strings that end with "at"
str_view(x, pattern = "at" %R% END)
```

***Instruction 4:***

Provide the pattern argument to match strings that are `"cat"` exactly. Think about this as matching the start of the string, followed by `"cat"` and then the end of the string.

```{r}
# Match the strings that is exactly "cat"
str_view(x, pattern = START %R% "cat" %R% END)
```

##### 3.3 Matching any character 

In a regular expression you can use a wildcard to match a single character, no matter what the character is. In `rebus` it is specified with `ANY_CHAR`. Try typing `ANY_CHAR` in the console. You should see that in the regular expression language this is specified by a dot, `.`.

For example, `"c" %R% ANY_CHAR %R% "t"` will look for patterns like "c_t" where the blank can be any character. Consider the strings: `"cat"`, `"coat"`, `"scotland"` and `"tic toc"`. Where would the matches to `"c" %R% ANY_CHAR %R% "t"` be?

Test your intuition by running:

>str_view(c("cat", "coat", "scotland", "tic toc"), 
  pattern = "c" %R% ANY_CHAR %R% "t")
  
Notice that `ANY_CHAR` will match a space character (`c t` in `tic toc`). It will also match numbers or punctuation symbols, but `ANY_CHAR` will only ever match one character, which is why we get no match in `coat`.

***Instruction 1:***

Match any character followed by a `t`.

```{r}
# Match two characters, where the second is a "t"
str_view(x, pattern = ANY_CHAR %R% "t")
```

***Instruction 2:***

Match a `t` followed by any character. Notice how the final `t` in `cat` and `coat` don't match, that's because there is no character after the `t` to match to `ANY_CHAR`.

```{r}
# Match a "t" followed by any character
str_view(x, pattern = "t" %R% ANY_CHAR)
```

***Instruction 3:***

Match any two characters. Notice the first two characters are matched. Regular expressions are lazy and will take the first match they find.

```{r}
# Match two characters
str_view(x, pattern = ANY_CHAR %R% ANY_CHAR)
```

***Instruction 4:***

Match a string that is exactly three characters.

```{r}
# Match a string with exactly three characters
str_view(x, pattern = START %R% ANY_CHAR %R% ANY_CHAR %R% ANY_CHAR %R% END)
```

##### 3.4 Combining with stringr functions 

You can pass a regular expression as the `pattern` argument to any `stringr` function that has the `pattern` argument. You can use `str_detect()` to get a logical vector for whether there was a match, `str_subset()` to return just the strings with matches, and `str_count()` to count the number of matches in each string.

As a reminder, compare the output of those three functions with our "c_t" pattern from the previous exercise:

>x <- c("cat", "coat", "scotland", "tic toc")
pattern <- "c" %R% ANY_CHAR %R% "t"
str_detect(x, pattern)
str_subset(x, pattern)
str_count(x, pattern)

It now also makes sense to add `str_extract()` to your repertoire. It returns just the part of the string that matched the pattern:

>str_extract(x, pattern)

You'll combine your regular expression skills with `stringr` to ask how often a `q` is followed by any character in boy names.

It's always a good idea to test your pattern, so this pattern is shown matched with four names. The first two shouldn't have matches (can you explain why?) but the last two should.

***Instruction 1:***

- Find the `boy_names` with the pattern by using `str_subset()`. Assign the result to `names_with_q`.

- Run `length()` on the result to find out how many there are.

```{r echo=FALSE}
library(tidyverse)
library(babynames)
babynames_2014 <- filter(babynames, year == 2014)
boy_names <- filter(babynames_2014, sex == "M")$name
girl_names<- filter(babynames_2014, sex == "F")$name
```

```{r}
pattern <- "q" %R% ANY_CHAR

# Find names that have the pattern
names_with_q <- str_subset(boy_names, pattern)

# How many names were there?
length(names_with_q)
```

***Instruction 2:***

- Find just the part of `boy_names` that matched with `str_extract()`. Assign the result to `part_with_q`.

- Run `table()` on the result to find out how many have `qu` and how many have other patterns.

```{r}
# Find part of name that matches pattern
part_with_q <- str_extract(boy_names, pattern)

# Get a table of counts
table(part_with_q)
```

***Instruction 3:***

- Check that there weren't any `boy_names` that might have had the pattern twice (you would have only found the first match) by using `str_count()`. Assign the result to `count_of_q`.

- Use `table()` on the result.

```{r}
# Did any names have the pattern more than once?
count_of_q <- str_count(boy_names, pattern)

# Get a table of counts
table(count_of_q)
```

***Instruction 4:***

- Get a logical vector of whether or not each boy's name contains `q*` by calling `str_detect()`. Assign the result to `with_q`.

- Count the fraction of boy's names containing `q*` by calculating the `mean()` of `with_q`.

```{r}
# Which babies got these names? (get logical vector back)
with_q <- str_detect(boy_names, pattern)

# What fraction of babies got these names? (get mean)
mean(with_q)
```

##### 3.5 More regular expressions (video)

##### 3.6 Alternation 

The `rebus` function `or()` allows us to specify a set of alternatives, which may be single characters or character strings, to be matched. Each alternative is passed as a separate argument.

For example, `or("grey", "gray")` allows us to detect either the American or British spelling:

>x <- c("grey sky", "gray elephant")
str_view(x, pattern = or("grey", "gray"))

Since these two words only differ by one character you could equivalently specify this match with `"gr" %R% or("e", "a") %R% "y"`, that is "a `gr` followed by, an `e` or an `a`, then a `y`".

Notice we've added the argument `match = TRUE` to `str_view()`, this will only display elements that had a match, which is useful when you are searching over many strings.

***Instruction 1:***

Create a pattern that matches `Jeffrey` or `Geoffrey` in `boy_names`. Assign this to `whole_names`.

```{r}
# Match Jeffrey or Geoffrey
whole_names <- or("Jeffrey", "Geoffrey")
str_view(boy_names, pattern = whole_names, match = TRUE) 
```

***Instruction 2:***

Create a pattern that matches `Je` or `Geo` followed by `ffrey`. Assign this to `common_ending`. You should get the same matches as the previous pattern.

```{r}
# Match Jeffrey or Geoffrey, another way
common_ending <- or("Je", "Geo") %R% "ffrey"
str_view(boy_names, pattern = common_ending, match = TRUE)
```

***Instruction 3:***

Create a pattern that matches `Je` or `Geo`, followed by `ff`, followed by one of `ry`, `ery` `rey`, or `erey`. Assign this to `by_parts`.

```{r}
# Match with alternate endings
by_parts <- or("Je", "Geo") %R% "ff" %R% or("ry", "ery", "rey", "erey")
str_view(boy_names, pattern = by_parts, match = TRUE)
```

***Instruction 4:***

Create a pattern that matches all `girl_names` that start with `C` or `K` followed by `ath`. Assign this to `ckath`.

```{r}
# Match names that start with Cath or Kath
ckath <- or("C", "K") %R% "ath"
str_view(girl_names, pattern = ckath, match = TRUE)
```

##### 3.7 Character classes 

In regular expressions a character class is a way of specifying "match one (and only one) of the following characters". In `rebus` you can specify the set of allowable characters using the function `()`.

This is another way you could specify an alternate spelling, for example, specifying "a `gr` followed by, either an `a` or `e`, followed by a `y`":

>x <- c("grey sky", "gray elephant")
str_view(x, pattern = "gr" %R% ("ae") %R% "y")

A negated character class matches "any single character that isn't one of the following", and in `rebus` is specified with `negated_()`.

Unlike in other places in a regular expression you don't need to escape characters that might otherwise have a special meaning inside character classes. If you want to match `.` you can include `.` directly, e.g. `(".")`. Matching a `-` is a bit trickier. If you need to do it, just make sure it comes first in the character class.

***Instruction 1:***

- Create a character class that contains vowels, `a`, `e`, `i`, `o`, `u` and their upper case versions.

- Print `vowels`. In the regular expression language, a character class is put inside `[`.

- View the matches to the pattern `vowels` in `x` with `str_view()`. Notice how only the first vowel is matched.

```{r}
# Create character class containing vowels
vowels <- ("aeiouAEIOU")

# Print vowels
vowels

# See vowels in x with str_view()
str_view(x, vowels)
```

***Instruction 2:***

View the matches to the pattern `vowels` in `x` with `str_view_all()`. Now all matches are highlighted.

```{r}
# See vowels in x with str_view_all()
str_view_all(x, vowels)
```

***Instruction 3:***

- Find the number of vowels in each `boy_names` by combining `str_count()` with the `vowels` pattern.

- Find the number of characters in each `boy_names` with `str_length()`.

```{r}
# Number of vowels in boy_names
num_vowels <-  str_count(boy_names, vowels)

# Number of characters in boy_names
name_length <- str_length(boy_names)
```

***Instruction 4:***

- Find the average number of vowels in `boy_names` using the mean of `num_vowels`.

- Look at the mean ratio of `num_vowels` and `name_length`.

```{r}
# Calc mean number of vowels
mean(num_vowels)

# Calc mean fraction of vowels per name
mean(num_vowels/name_length)
```

##### 3.8 Repetition 

The `rebus` functions `one_or_more()`, `zero_or_more()` and `optional()` can be used to wrap parts of a regular expression to allow a pattern to match a variable number of times.

Take our `vowels` pattern from the last exercise. You can pass it to `one_or_more()` to create the pattern that matches "one or more vowels". Take a look with these interjections:

>x <- c("ow", "ooh", "yeeeah!", "shh")
str_view(x, pattern = one_or_more(vowels))

You'll see we can match the single `o` in `ow`, the double `o` in `ooh` and the string of `e`s followed by the `a` in `yeeeah`, but nothing in `shh` because there isn't a single vowel.

In contrast `zero_or_more()` will match even if there isn't an occurrence, try

> str_view(x, pattern = zero_or_more(vowels))

Since both `yeeeah` and `shh` start without a vowel, they match "zero vowels", and since regular expressions are lazy, they look no further and return the start of the string as a match.

***Instruction 1:***

Find `boy_names` that are only vowels, by combining `one_or_more()` with `vowels`. You will need to either specify `START` and `END` or use `exactly()`.

```{r}
# Vowels from last exercise
vowels <- ("aeiouAEIOU")

# See names with only vowels
str_view(boy_names, 
  pattern = exactly(one_or_more(vowels)), 
  match = TRUE)
```

***Instruction 2:***

- Create a `negated_char_class()` that matches anything but a vowel. (Note: `negated_char_class(vowels)` is not the right answer!)

- Find `boy_names` that have no vowels, by combining `exactly()` and `one_or_more()` with `not_vowels`.

```{r}
# Use `negated_()` for everything but vowels
not_vowels <- negated_char_class("aeiouAEIOU")

# See names with no vowels
str_view(boy_names, 
  pattern = exactly(one_or_more(not_vowels)), 
  match = TRUE)
```

##### 3.9 Shortcuts (video)

**Ranges in Character classes**

- A digit `char_class("0-9")`

- A low case letter `char_class("a-z")`

- An upper case letter `char_class("A-Z")`

- DGT: any digit, 0-9

- WRD: a word character

- SPC

##### 3.10 Hunting for phone numbers 

For your first task you are going to pull out the phone numbers from this vector of contact information:

>contact

You'll assume the phone numbers you are looking for follow the American standard of a three digit area code, a three digit exchange and then a four digit number, but each part could be separated by spaces or various punctuation symbols.

***Instruction 1:***

Phone numbers start with three digits. Create a pattern that matches three digits in a row. Use `str_view_all()` to test your pattern.

```{r echo=FALSE}
narratives <- readRDS("narratives.rds")

contact <- c("Call me at 555-555-0191", 
             "123 Main St", 
             "(555) 555 0191", 
             "Phone: 555.555.0191 Mobile: 555.555.0192")
```

```{r}
# Create a three digit pattern and test
three_digits <- DGT %R% DGT %R% DGT

# Test it
str_view_all(contact, pattern = three_digits)
```

***Instruction 2:***

There might be a range of separators, so make a `separator` pattern, that uses `()` to combine a character class containing `-`, `.`, `(`, `)`, and ` ` .

```{r}
# Create a separator pattern and test
separator <-  ("-.() ")

# Test it
str_view_all(contact, pattern = separator)
```

***Instruction 3:***

Put together a pattern to match a whole phone number: an optional open parenthesis (`OPEN_PAREN`), followed by three digits, followed by zero or more separators, followed by three digits followed by zero or more separators, followed by four digits.

```{r}
# Use these components
three_digits <- DGT %R% DGT %R% DGT
four_digits <- three_digits %R% DGT
separator <- ("-.() ")

# Create phone pattern
phone_pattern <- optional(OPEN_PAREN) %R%
  three_digits %R%
  zero_or_more(separator) %R%
  three_digits %R% 
  zero_or_more(separator) %R%
  four_digits
      
# Test pattern           
str_view_all(contact, phone_pattern)
```

***Instruction 4:***

- Extract the phone number from `contact`, using `str_extract`.

- Try using `str_extract_all()` instead. Can you see the difference?

```{r}
# Use this pattern
three_digits <- DGT %R% DGT %R% DGT
four_digits <- three_digits %R% DGT
separator <- ("-.() ")
phone_pattern <- optional(OPEN_PAREN) %R% 
  three_digits %R% 
  zero_or_more(separator) %R% 
  three_digits %R% 
  zero_or_more(separator) %R%
  four_digits
  
# Extract phone numbers
str_extract(contact, phone_pattern)

# Extract ALL phone numbers
str_extract_all(contact, phone_pattern)
```

##### 3.11 Extracting age and gender from accident narratives

Recall from the video, you want to parse out age and gender from accident narratives. For example, this narrative

> 19YOM-SHOULDER STRAIN-WAS TACKLED WHILE PLAYING FOOTBALL W/ FRIENDS 

describes male of age 19, and this one

>TRIPPED OVER CAT AND LANDED ON HARDWOOD FLOOR. LACERATION ELBOW, LEFT. 33 YOF*

a female of age 33.

You are generally looking for a pattern with a number, something to indicate the units, e.g. `YO` or `YR` for years old, or `MO` for months old, and a character that identifies the gender.

In this exercise you'll build up a pattern to pull out the part of the narrative that has the age and gender information. Then, in the next exercise you'll parse out the age and gender into separate variables.

***Instruction 1:***

Create an `age` pattern that matches one or two digits. Test your pattern using `str_view()`.

```{r}
# Pattern to match one or two digits
age <- DGT %R% optional(DGT)

# Test it
str_view(narratives, pattern = age)
```

***Instruction 2:***

- Create a `unit` pattern that matches an optional space, then one of `YO`, `YR` or `MO`.

- Check your pattern so far by using `str_view()` with the pattern `age %R% unit`.

```{r}
# Use this pattern
age <- DGT %R% optional(DGT)

# Pattern to match units 
unit <- optional(SPC) %R% or("YO", "YR", "MO")

# Test pattern with age then units
str_view(narratives, 
         pattern = age %R% unit)
```

***Instruction 3:***

- Create a `gender` pattern that matches an optional space then `M` or `F`.

- Check your pattern so far by using `str_view()` with the pattern `age %R% unit %R% gender`.

```{r}
# Use these patterns
age <- DGT %R% optional(DGT)
unit <- optional(SPC) %R% or("YO", "YR", "MO")

# Pattern to match gender
gender <- optional(SPC) %R% ("MF")

# Test pattern with age then units then gender
str_view(narratives, 
         pattern = age %R% unit %R% gender)
```

***Instruction 4:***

Extract the age-unit-gender piece from each narrative using `str_extract()`.

```{r}
# Use these patterns
age <- DGT %R% optional(DGT)
unit <- optional(SPC) %R% or("YO", "YR", "MO")
gender <- optional(SPC) %R% or("M", "F")

# Extract age_gender, take a look
age_gender <- str_extract(narratives, pattern = age %R% unit %R% gender)
age_gender
```

##### 3.12 Parsing age and gender into pieces 

To finish up, you need to pull out the individual pieces and tidy them into usable variables.

There are a few ways you could get at one piece: you could extract out the piece you need, you could replace everything that isn't the piece you need with `""`, or you could try to split into the pieces you need. You'll try a few of these in this exercise and you'll see yet another way in the next chapter. For the first option, `stringr` has a nice convenience function, `str_remove()`, that works like `str_replace()` with `replacement = ""`.

One benefit of building up your pattern in pieces is you already have patterns for each part that you can reuse now.

***Instruction 1:***

Use `str_extract` with your `age` pattern to extract just the age from `age_gender`, then transform it to a number with `as.numeric()`.

```{r}
# age_gender, age, gender, unit are pre-defined
ls.str()

# Extract age and make numeric
as.numeric(str_extract(age_gender, age))
```

***Instruction 2:***

- Create `genders` by using `str_remove()` with your `age %R% unit` pattern to replace everything except the gender with `""`.

- `genders` has a few extra spaces; remove them.

```{r}
# Replace age and units with ""
genders <- str_remove(age_gender, pattern = age %R% unit)

# Replace extra spaces
str_remove_all(genders, pattern = one_or_more(SPC))
```

***Instruction 3:***

- Get `time_units` by using `str_extract()` on `age_gender` with your `unit` pattern.

- To know if the units are months or years we just need the first character after any spaces. Use `str_extract()` on `time_units` with the pattern `WRD` to get `time_units_clean`.

- Complete the final line to convert any ages reported in months to an age in years.

```{r}
# Numeric ages, from previous step
ages_numeric <- as.numeric(str_extract(age_gender, age))

# Extract units 
time_units <- str_extract(age_gender, unit)

# Extract first word character
time_units_clean <- str_extract(time_units, WRD)

# Turn ages in months to years
ifelse(time_units_clean == "Y", ages_numeric, ages_numeric / 12)
```

